---
title: "Sierra Amphibian Abundance"
author: "Kerry Nixon"
date: "1/23/2021"
output: html_document
---

```{r setup, include=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(janitor)
library(ggrepel)
library(ggExtra)
library(readxl)
library(gghighlight) # For useful highlighting
library(lubridate) # For dates & times
library(plotly) # Interactive graphs
library(patchwork) # For compound figures
```


```{r}
# Importing data

## read in data - data is tidy and headers are already in lower_snake_case
sierra_amphibians <- read_excel(here("data", "sierra_amphibians.xlsx"), na = "NA")

```

### Part 1 - Plotting Abundance of Ramu Muscosa by Year and Life Stage
```{r, message=FALSE, warning=FALSE}
# Graph 1: Wrangling Yellow-Legged Mountain frogs, updating date format, pulling year

## subset - filtering for Rana muscosa "RAMU" species and excluding "EggMass" life stage
ra_mu <- sierra_amphibians %>% 
  filter(amphibian_species == "RAMU",
         amphibian_life_stage != "EggMass") %>% 
  mutate(survey_date = ymd(survey_date)) %>% 
  mutate(year = year(survey_date)) %>% 
  drop_na(amphibian_number)

## saving summary of Yellow-Legged Mountain Frog counts by life stage and year subset
stg_yr_sum <- ra_mu %>% 
  group_by(year, amphibian_life_stage) %>% 
  summarise(count = sum(amphibian_number))
```


```{r}
# Graph 1: DataViz - Creating graph of counts from summary

## Dataviz Exploration: trying out different visualizations for this subset
viz_1 <- ggplot(data = stg_yr_sum, aes(fill = amphibian_life_stage, y = amphibian_life_stage, x = count)) +
  geom_bar(position = "dodge", stat = "identity") +
  facet_wrap(~year, scales = "free_x")

viz_2 <- ggplot(data = stg_yr_sum, aes(x = year, y = count)) +
  geom_line(aes(color = amphibian_life_stage)) +
  geom_point(size = 0.5) +
  scale_x_continuous(n.breaks = 10) +
  theme_bw()

## This is the better of the graphs for the summary subset, clearly showing when data was and wasn't collected, and cumulative as well as group values
n_by_year <- ggplot(data = stg_yr_sum, aes(fill = amphibian_life_stage, x = year, y = count)) +
  geom_bar(position = "stack", stat = "identity") +
  scale_x_continuous(n.breaks = 10) + 
  scale_y_continuous(breaks = seq(0,185000,25000), expand = expansion(mult = c(0, .1))) +
  scale_fill_manual(values = c("#76448A","#D4AC0D", "#1F618D")) +
  labs(x = "\nYear\n",
       y = "\nNumber of Specimens Counted\n",
       fill = "Life Stage") +
  theme_bw() +
  removeGridX() +
  theme(legend.position = c(0.2,0.7),
        legend.background = element_rect(fill = "white",
                                         size = 0.5,
                                         linetype = "solid",
                                         color = "black"))

# Visual stored as object for use in compound figure as instructed in assignment
```

### Part 2 - Plotting Sierra Lakes with Highest Cumulative Rana Muscosa Counts
```{r, message=FALSE, warning=FALSE}
# Graph 2: Wrangling for top 5 lakes with highest YLMF counts over study period

## subset - removing "Tadpole" life stage, summarizing count by lake id
top_5 <- ra_mu %>% 
  filter(amphibian_life_stage != "Tadpole") %>% 
  mutate(lake_id_lbl = sub("^", "Lake ", lake_id)) %>% 
  drop_na(amphibian_number) %>% 
  group_by(lake_id_lbl) %>% 
  summarise(count = sum(amphibian_number)) %>% 
  slice_max(count, n = 5)
```

```{r}
# Graph 2: DataViz

## Creating a column graph of 5 lakes with highest Rana Muscosa counts
top_5_lakes <- ggplot(data = top_5, aes(x = fct_reorder(lake_id_lbl, -count), y = count)) +
  geom_col(fill = "#85C1E9",
           show.legend = FALSE) +
  scale_y_continuous(expand = expansion(mult = c(0, .1))) +
  labs(x = "\nLake Site Number\n",
       y = "\nNumber of Specimens Counted\n") +
    theme_bw()

# Visual stored as object for use in compound figure as instructed in assignment
```

### Mountain Yellow-Legged Frog Abundance in the Southern Sierra Nevada
```{r, fig.width = 12}
# Putting the two plots together into a compound figure with `patchwork`

cmpnd_fig <- (n_by_year | top_5_lakes) +
  plot_annotation(tag_levels = "A", tag_suffix = ".")

## calling the compound figure
cmpnd_fig
```
**Figure 1: A.** This stacked bar graph shows the number of mountain yellow-legged frogs (*Rana Muscosa*) counted each year between 1995-2002, at over 8,000 stationary water bodies in the southern Sierra Nevada, including lakes, ponds, marshes, and meadows. Observations were not recorded for 1998 or 1999. In 1995 the count of frogs at each of the three life stages was so small that the bar representing that count is barely visible above the x-axis. **B.** This bar plot shows the five lakes with the highest cumulative counts of adult and sub-adult mountain yellow-legged frogs over the course of the study (1999-2002). Tadpoles were not included in total counts. **Data source:** Sierra Lakes Inventory Project


**Data citation:** Knapp, R.A., C. Pavelka, E.E. Hegeman, and T.C. Smith. 2020. The Sierra Lakes Inventory Project: Non-Native fish and community composition of lakes and ponds in the Sierra Nevada, California ver 2. Environmental Data Initiative. https://doi.org/10.6073/pasta/d835832d7fd00d9e4466e44eea87fab3

